---
title: "09-2 Problem Set"
format: pdf
---

Complete the following code chunks and replacing the *answer here* text. Knit this document to a pdf, **double check the pdf document for errors**, and submit this to this week's Problem Set assignment on Canvas. 

## Part 1. 

Use the dataset from the paper: *Statistical tools for water quality assessment and monitoring in river ecosystems – a scoping review and recommendations for data analysis* to assess the whether water_depth and total concentration of vanadium interact to predict the dissolved concentration of vanadium. Note that total vs dissolved concentrations are in the type column.  

```{r}
library(tidyverse)
metals <- read_csv("data/S1_data_cleaned.csv")
```

A) Filter out just the vanadium data, and create dissolved and total concentration variables (from the existing `type` and `concentration` variables) using `pivot_wider`.

```{r}

```

B) Make an ANOVA model of how the full set of interactions between `water_depth`, `station_id`, and `total` (vanadium) is predictive of `dissolved` (vanadium). 

```{r}

```


C) What variables and interactions are significant in your model?
*answer here*


D) What variables explains the most variation in your model?
*answer here*


E) What other variables might be added to improve your model? Why?
*answer here*


F) Does adding `yr` to your model improve it (in terms of `Sum Sq` variation explained by variables vs residuals)?
```{r}

```

G) Use `anova` to compare the two models.

```{r}

```

H) Explain the results and which model is a better fit of the data.

## Part 2. 

Use non-linear least squares to model the inhibitor data from last week. Using the following equation: 
$$C_{bacteria} = \phi_1 e^{\phi_2 t}$$

We could make this into a linear form by taking the natural log of both sides:

$$\ln C_{bacteria} = \ln(\phi_1) + \phi_2 t$$

The concentration of bacteria in a batch reactor liquid culture is monitored after adding an inhibitor, yielding the following data:

time [hours] <- 4 8 12 16 20 24
c_bacteria [CFU/mL] <- 1600 1320 1000 890 650 560

where "CFU" is  colony forming units, or how many colonies grow on a petri dish when putting a known volume of culture on the solid agar media in the petri dish. Each colony grows from a single viable (alive) cell. 

A) Fit a linearized model of the above equation to the data (after transformation).

B) Fit a non-linear model to the data. 

```{r}

```

C) Compare how this non-linear model improves the prediction of the concentration at time = 0 versus a linear model (particularly the standard error in the y-intercept for the linearized equation model, vs $\phi_1$ in the non-linear model). Use propagation of uncertainty for the linear model (<https://en.wikipedia.org/wiki/Propagation_of_uncertainty>). In this and most cases the standard error in the parameters is equivalent to a standard deviation in measurements.
 
```{r}


```

For this linear model the value of $\phi_1$ is related to the y-intercept, $y_0$ as $e^{y_0}$. So to calculate the value of $\phi_1$:
$$\phi_1 = e^{y_0}$$. 

Looking at the table in wikipedia for similar calculations we find:
$$f = ae^{bA}$$ 
where here $a$ and $b$ are 1 and therefore $A = y_0$ and $f = \phi_1$. 
The standard deviation in $f$, $\sigma_f$, from the third column in the table is
$$\sigma_f = |f| |b\sigma_A|$$ 
If we translate this to the terminology of our specific problem we get:
$$\sigma_{\phi_1} = |\phi_1| |\sigma_{y_0}|$$ 